### 2020-01-30 Mathematical Statistics/05_conf_int_bootstrap.rmd
bootstrap
bootstrap

icecream <- IceCream %>%
   mutate(Vanilla = VanillaCalories,
          Chocolate = ChocolateCalories) %>%
   select(Vanilla, Chocolate) %>%
   melt(variable.name = "Flavor", value.name = "Calories") %>%
   as.data.table()

@@

@@

@@

>>>>

```{r}
icecream <- IceCream %>%
   mutate(Vanilla = VanillaCalories,
          Chocolate = ChocolateCalories) %>%
   select(Vanilla, Chocolate) %>%
   melt(variable.name = "Flavor", value.name = "Calories") %>%
   as.data.table()

@@

```{r}
Diving2017 <- data.table(read.csv(paste0(data.dir, "Diving2017.csv"),
                                 header = T))
```

a.) Compute the difference in the median scores in the final and semi-final rounds.

```{r}
observed <- median(Diving2017$Final) - median(Diving2017$Semifinal)
```

b.) Calculate a 95% bootstrap interval for this statistic.

```{r}
N <- 10^5

result <- vector(mode = "numeric", length = N)

for(i in 1:N)
{
   index <- sample(nrow(Diving2017), replace = T)

   Dive.boot <- Diving2017[index, ] # resample pairs
   result[i] <- median(Dive.boot$Final) - median(Dive.boot$Semifinal)
}

ggplot(data.table(values = result), aes(values)) +
   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
   geom_density(col = "darkorange")
```

### 5.21

Two college students collecteed data on the price of hardcover textbooks from two disciplinary areas: Mathematics and the Natural Sciences and the Social Sciences. The data are in the file BookPrices.

```{r}
BookPrices <- data.table(read.csv(paste0(data.dir, "BookPrices.csv"),
                                 header = T))
```

a.) Perform some exploratory data analysis on book prices for each of the two diciplinary areas.

```{r}
ggpairs(BookPrices)

ggplot(BookPrices, aes(Area, Price, group = Area)) +
   geom_boxplot(aes(fill = Area))
```

b.) Bootstrap the mean of book price for each area separately, and describe the distributions.

@@

prices.math <- BookPrices[Area == "Math & Science"]$Price
prices.science <- BookPrices[Area == "Social Sciences"]$Price

N <- 10e4

bootstrap.values <- vector(mode = "numeric", length = N); n <- length(prices.math)
observed <- mean(prices.math)

for(i in 1:N)
{
   bootstrap.values[i] <- mean( sample(prices.math, n, replace = T) )
}

bootstrap.mean <- mean(bootstrap.values)

@@

```{r}
N <- 10e4

bootstrap.values <- vector(mode = "numeric", length = N); n <- length(prices.science)
observed <- mean(prices.science)

for(i in 1:N)
{
   bootstrap.values[i] <- mean( sample(prices.science, n, replace = T) )
}

bootstrap.mean <- mean(bootstrap.values)
```

_Both of the distributions are relatively symetrical and approximately normal._

c.) Bootstrap the ratio of means. Provide plots of the bootstrap distribution and comment.

```{r}
N <- 10e4

bootstrap.values <- vector(mode = "numeric", length = N); n <- length(prices.science)
observed <- mean(prices.math) / mean(prices.science)

for(i in 1:N)
{
   samp.math <- sample(prices.math, length(prices.math), replace = T)
   samp.science <- sample(prices.science, length(prices.science), replace = T)

   bootstrap.values[i] <- mean(samp.math) / mean(samp.science)
}

bootstrap.mean <- mean(bootstrap.values)
```

_The ratio of means distribution is heavly skewed with a long right tail._

d.) Find the 95% bootstrap percentile interval for the ratio of means. Interpret this interval.

```{r}
alpha <- 0.05

_There is statistically significant evidence that the cost of math books are higher than science books._

e.) What is the bootstrap estimate of the bias? What fraction of the bootstrap standard error does it represent?

```{r}
bootstrap.bias <- mean(bootstrap.values) - observed
bootstrap.se <- sd(bootstrap.values)

bootstrap.bias

The bias in the bootstrap is quite high at 5%. Also, the bias is almost 16% of the SE which is extremely high.

### 5.22

Import the data from flight delays case study in S1.1.

For this study, we will consider the ratio of means.

```{r}
FlightDelays <- data.table(read.csv(paste0(data.dir, "FlightDelays.csv"),
                                 header = T))
```

a.) Perform some exploratory data analysis on flight delay lengths.

```{r}
ggpairs(FlightDelays)

ggplot(FlightDelays, aes(Carrier, Delay, col = Carrier)) +
   geom_boxplot()
```

b.) Bootstrap the mean of flight delay lengths for each airline separately, and describe the distribution.

```{r}
ua.delay <- FlightDelays[Carrier == "UA"]$Delay
aa.delay <- FlightDelays[Carrier == "AA"]$Delay

n <- length(ua.delay); N <- 10e4
bootstrap.values <- vector(mode = "numeric", length = n)

for(i in 1:N)
{
   bootstrap.values[i] <- mean( sample(ua.delay, n, replace = T))   
}

ggplot(data.table(values = bootstrap.values), aes(values)) +
   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
   geom_density(col = "darkorange")

n <- length(aa.delay); N <- 10e4
bootstrap.values <- vector(mode = "numeric", length = n)

for(i in 1:N)
{
   bootstrap.values[i] <- mean( sample(aa.delay, n, replace = T))   
}

ggplot(data.table(values = bootstrap.values), aes(values)) +
   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
   geom_density(col = "darkorange")

c.) Bootstrap the ratio of means. Provide plots of the bootstrap distribution and describe the distribution.

```{r}
n <- length(aa.delay); N <- 10e4
bootstrap.values <- vector(mode = "numeric", length = n)

observed <- mean(aa.delay) / mean(ua.delay)

for(i in 1:N)
{
   bootstrap.values[i] <- mean( sample(aa.delay, n, replace = T)) / mean( sample(ua.delay, n, replace = T))   
}

ggplot(data.table(values = bootstrap.values), aes(values)) +
   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
   geom_density(col = "darkorange") +
   geom_vline(xintercept = observed)
```

d.) Find and interpret the 95% confidence bootstrap interval for the ratio of means.

Distribution is approximately normal, and the interval contains zero so we can't rule out chance in the flight delay data.

e.) What is the bootstrap estimate of the bias? What fraction of the standard error does it represent?

```{r}
boot.bias <- mean(bootstrap.values) - observed

boot.bias / sd(bootstrap.values)
```

The se is approximately 3% of the bias.
